---
title: "Kashyap_HW5"
author: "Madhuri"
date: "`r Sys.Date()`"
output: html_document
---

## Homework 5

```{r setup, include=FALSE}
library(janitor)
library(tidymodels)
library(tidyverse)
library(ggplot2)

df <- read.csv("Pokemon.csv")
```


### Question 1

```{r}
df_clean <- as_tibble(df) %>% 
  clean_names()
```

The clean_names() function altered the column names, making them more readable This has obvious benefits in increasing usability by making all the characters lower case and removing any special symbols.


### Question 2

```{r}
#barplot
ggplot(df_clean, aes(x = type_1)) +
  geom_bar(stat = "count", fill = "steelblue") +
  theme_classic() +
  coord_flip() +
  xlab("Primary Type") +
  ylab("Count") 

#number of classes =18
summary(factor(df_clean$type_1))

#filtering data
df_filtered <- filter(df_clean, type_1 == "Bug" | type_1 == "Fire" | type_1 == "Grass" |
                        type_1 == "Normal" | type_1 == "Water" | type_1 == "Psychic")
#checking new dataframe
summary(factor(df_filtered$type_1))

#converting to factors
df_filtered$type_1 <- as.factor(df_filtered$type_1)
df_filtered$legendary <- as.factor(df_filtered$legendary)
```


There are 18 classes of primary types of Pokemon.
The Flying an Fairy classes have the lowest number of Pokemon with 4 and 17, respectively.


### Question 3

```{r}
set.seed(1212)
df_split <- initial_split(df_filtered, strata = "type_1")

df_train <- training(df_split)
df_test <- testing(df_split)

df_fold <- vfold_cv(df_train, v = 5, strata = "type_1")
```


Stratified k-fold cross=validation is particularly useful in cases like our where the variable of interest (type_1) is imbalanced, implying that the frequencies of the classes withing that variable are not equally distributed. This type of cross-validation then allows us to ensure that each of the folds have a similar distribution of data withing type_1 as the original dataframe.
